---
title: "Add All Relationships to Master File Table"
date: "Created: 2018-08-29 <br> Updated: `r Sys.Date()`"
output: 
  html_notebook:
    css: custom-css.css
---

# Table of contents

1. [Overview](#overview)   
2. [User-created functions](#functions)   
3. [Add file relationships to master file table](#relationship)   
4. [](#)     


# Overview {#overview}

I needed to come up with a better method for tracking notes/documentation/file dependencies in analysis projects. The goal of Pathfinder is to help track file dependencies in research projects created in RStudio. Perhaps other things later. 

Assumptions:    
1. Pathfinder will typically be deployed in R projects.   

  * When deployed in R projects, Pathfinder will automatically detect the .Rproj file and the path to that .Rproj file will be se as the root directory.   

  * If the user desires a different behavior (e.g., target a different project than the one the user is currently working in or mapping dependencies among a collection of files that do not exist in an R project) then the path can be specified manually.    

    - The specified path will serve as the root directory. Pathfinder will attempt to gather dependency information about every file in the root directory and all child directories.   
        
    - If the user is not in an R project and a path is not specified, then Pathfinder will set the current directory to be the project root. Pathfinder will attempt to gather dependency information about every file in the root directory and all child directories.
        
2. Therefore, to use Pathfinder the user must be in an open R project, provide a file path, or both.


# Load packages and data

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r message=FALSE}
library(tidyverse)
```

## Root directory path

```{r}
root_dir <- readr::read_rds("data/root_dir.rds")
```

## List of files under the root directory

```{r}
root_dir_file_list <- readr::read_rds("data/root_dir_file_list")
```

## Master file table

```{r}
root_dir_file_table <- readr::read_rds("data/root_dir_file_table.rds")
```











# User-created functions {#functions}

These are the functions we use below. I'm making them easy to find here at the top of the notebook because you may want to move them over to the Pathfinder package when you are done with testing.

## Function 1

Description

```{r}

```

```{r}

```


## Save

For use in future

```{r}
# readr::write_rds(root_dir_file_table, "data/root_dir_file_table.rds")
```

[top](#top)










# Add file relationships to master file table {#relationship}

What I eventually need is a table like the one on line 76 of make_edges.Rmd. Specifically, I need to have the following columns in the master file table:

* reads in (already done)   
* creates (done for data, but not for output)   
* is read into   
* is created in   

Need to figure out an automated way to do that.

* Go back to 04_look_for_k

```{r}
root_dir_file_table
  
```


[top](#top)

&nbsp;

-------------------------------------------------------------------------------

```{r echo=FALSE}
sessionInfo()
```
